--- 
title "simulation and the Central limit theorem: statistical inference course project"
author: "Peter Thompson"
---
# Inference

## Synopsis
something about toothgrowth


### Data Summary

 |supp | dose | meanlen | varlen |
 |-----|------|---------|--------|
 |  OJ |  0.5 |  13.23  |19.8890 |
 |  OJ |  1.0 |  22.70  |15.2956 |
 |  OJ |  2.0 |  26.06  | 7.0493 |
 |  VC |  0.5 |   7.98  | 7.5440 |
 |  VC |  1.0 |  16.77  | 6.3268 |
 |  VC |  2.0 |  26.14  |23.0182 |

OJ seems to do more.

hypothesis tests:
OJ is better than VC
higher dose is better
growth per dose?

growth per dose,  (or oj >< VC, two sided)
dosage - test 0.5 vs 2.0, 0.5 vs 1.0, 1.0 vs 2.0
Dosage regardless of method
method regardless of dose
method - growth per unit dose

t distribution assumes data iid normal "but it is robust" to this
unequal variance (the variances above are not equal)
https://github.com/bcaffo/courses/blob/master/06_StatisticalInference/08_tCIs/index.pdf
^ slide 21

Can't use paired test, because we can't compute the differences (we have 60 seperate hamsters, not 10 that were fed vitamin c multiple times via different methods). independent group t intervals (unequal variance), with alternative that the difference is not equal to zero (mu2.0 - mu 0.5)

## Appendix

```{r,load}
head(ToothGrowth)
ToothGrowth %>% group_by(supp,dose) %>% summarise(meanlen=mean(len),varlen=var(len))
```

```{r,quicktest}
vc <-ToothGrowth %>% filter(supp=='VC')
oj <-ToothGrowth %>% filter(supp=='OJ')
moose<-t.test(x=oj$len,y=vc$len,paired=F)
moose
```


```{r,plot}
ggplot(data=ToothGrowth,aes(x=len)) + geom_histogram(aes(fill=..count..)) + facet_grid( supp ~dose) + scale_fill_gradientn(colours=c('blue','purple'))
```

