--- 
title "simulation and inference: statistical inference course project"
author: "Peter Thompson"
---
# Introduction

This project is about simulation and inference and stuff



```{r}
summary(cars)
```


This summation expression $\sum_{i=1}^n X_i$ appears inline.


$$\sum_{i=1}^n X_i$$

# Simulation

```{r packages}
#opts_chunk$set(fig.width=8, fig.height=8,dpi=144)
require(ggplot2)
require(dplyr)
require(ggmap)
require(gridExtra)
```


```{r, generate}
set.seed(5074491)
lambda<-0.2
samples1k<-rexp(1000,rate=lambda)
means1kx40<-replicate(1000,mean(rexp(40,rate=lambda)))
data<-data.frame(samples1k,means1kx40)
names(data)<-c('expSamples','meanExp40Samples')

```

## sample mean

For random variables drawn from a (continuous) exponential distribution of rate $\lambda$, the propability dirstribution function is given by
$$
P(x) = \lambda e^{-\lambda x}.
$$ 

This distribution has a population mean of $\lambda$ and a standard deviation of $\sigma = \lambda$

```{r plot}
# plot distributions
hist_exp<-ggplot(data=data,aes(x=samples1k,fill=..count..)) + geom_histogram(binwidth=0.2) +
labs(title='distribution of 1000 random exponential variables',x='value',y='count') +
scale_fill_gradientn(colours=c('blue','purple'))
hist_expMean<-ggplot(data=data,aes(x=means1kx40,fill=..count..)) + geom_histogram(binwidth=0.2) +
labs(title='distribution of 1000 sample means of 40 random exponential variables',x='value',y='count') +scale_fill_gradientn(colours=c('blue','purple'))

grid.arrange(hist_exp,hist_expMean,nrow=1)
```



Part 1: Simulation Exercise Instructions
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the **properties of the distribution of the mean of 40 exponentials** . You should

Show the sample mean and compare it to the theoretical mean of the distribution.
Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
Show that the distribution is approximately normal.


```{r}


```
